\documentclass[conference]{IEEEtran}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{balance}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning, fit, backgrounds}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny,
  xleftmargin=2em,
  columns=fullflexible,
  keepspaces=true,
}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue,
}

\begin{document}

\title{VeilGuard: Preventing Credential Exposure\\Through AI Coding Assistants}

\author{
\IEEEauthorblockN{Manam Bharadwaj}
\IEEEauthorblockA{
\textit{Independent Researcher}\\
\url{https://github.com/manambharadwaj/veilguard}}
}

\maketitle

\begin{abstract}
AI coding assistants like Claude Code, Cursor, and GitHub Copilot routinely
read project files and store conversation transcripts on disk, opening up a
new avenue for credential exposure that existing secret scanners were not built
to handle. This paper introduces \textbf{VeilGuard}, an open-source Python
tool that tackles this problem on five fronts: regex scanning across 55
credential patterns in 8 categories, tool-aware hook and deny-rule installation
for 6 major AI assistants, deep transcript redaction, environment-variable
exposure checks, and an AES-256-GCM encrypted local secret store. We evaluate
VeilGuard on a curated benchmark of 93 files and pit it against three
widely used secret scanners. On our corpus, VeilGuard reaches perfect precision
and recall (F1~=~1.00) at 8{,}970 files per second, whereas gitleaks scores an
F1 of 0.85, trufflehog 0.53, and detect-secrets 0.00. The key differentiator
is that VeilGuard sits on the developer's machine and catches credentials
\emph{before} they enter the AI context window, rather than after they have
already been committed to version control.
\end{abstract}

\begin{IEEEkeywords}
secret detection, credential scanning, AI coding assistants, software security,
developer tools, LLM security
\end{IEEEkeywords}

% ============================================================
\section{Introduction}
\label{sec:intro}

AI coding assistants have quickly become part of everyday software development.
Tools like Claude Code~\cite{anthropic2024claude}, GitHub
Copilot~\cite{copilot2024}, Cursor~\cite{cursor2024}, Windsurf, Cline, and
Aider ingest source code, configuration files, and instruction documents to
offer context-aware completions, refactoring suggestions, and interactive
debugging. Many of these tools also record conversation history in local
transcript files and send project context to remote large language model (LLM)
endpoints.

This setup creates a problem that has received surprisingly little attention:
\textbf{credentials can leak through the AI context}. If an API key, a
database connection string, or a private key sits in a file the assistant reads,
that credential can end up in several unwanted places:

\begin{enumerate}
  \item On a remote LLM endpoint, where the model provider may log or cache it.
  \item In generated code or conversation responses the assistant echoes back.
  \item In local transcript files that other processes on the machine can read.
  \item In the hands of an attacker who has planted prompt-injection payloads
        inside the repository.
\end{enumerate}

Today's secret scanning tools~\cite{gitleaks2024,trufflehog2024,
detectsecrets2024} were designed to catch credentials in \emph{git history} or
at the \emph{CI pipeline} stage. They flag secrets that have already been
committed. None of them deal with the real-time risk of an AI assistant reading
credentials straight from the working tree, and none offer a way to scrub
credentials from transcript files or to stop the assistant from opening
secret-bearing files in the first place.

This paper presents \textbf{VeilGuard}, a Python CLI that provides layered
protection against credential exposure to AI coding assistants. VeilGuard runs
entirely on the developer's workstation, makes no network calls, and covers five
areas:

\begin{enumerate}
  \item \textbf{Detection.} 55 regex patterns across 8 credential categories
        scan source, configuration, and transcript files.
  \item \textbf{Prevention.} Tool-specific hooks, deny rules, and instruction
        files keep AI assistants away from secret-bearing files.
  \item \textbf{Remediation.} A deep transcript redaction engine replaces
        credentials in Claude Code JSONL transcripts with sanitized placeholders.
  \item \textbf{Verification.} A monitoring command ensures that
        environment-variable-backed secrets have not appeared in AI context files.
  \item \textbf{Isolation.} An AES-256-GCM encrypted local secret store offers
        a safe alternative to hard-coding credentials.
\end{enumerate}

Our contributions are as follows:

\begin{itemize}
  \item We define and formalize the threat model for credential exposure through
        AI coding assistants (Section~\ref{sec:threat}).
  \item We describe VeilGuard's design and implementation, which, to our
        knowledge, is the first tool built specifically for this threat class
        (Section~\ref{sec:design}).
  \item We evaluate VeilGuard on a curated benchmark corpus and show that it
        outperforms three established secret scanners in detection accuracy and
        throughput (Section~\ref{sec:eval}).
  \item We release VeilGuard as open-source software on PyPI and
        GitHub.\footnote{\url{https://github.com/manambharadwaj/veilguard}}
\end{itemize}


% ============================================================
\section{Background and Related Work}
\label{sec:related}

\subsection{Secret Scanning in Software Development}

Hard-coded credentials in source code continue to be a widespread security
issue~\cite{meli2019bad}. The community has developed several families of tools
to address it.

\textbf{Regex-based scanners.}
Gitleaks~\cite{gitleaks2024} and trufflehog~\cite{trufflehog2024} ship with
large pattern libraries for known credential formats and scan git repositories.
Gitleaks provides over 150 rules and works on git diffs. Trufflehog goes
further by adding entropy analysis and optional server-side verification of
detected secrets.

\textbf{Entropy-based detection.}
Detect-secrets~\cite{detectsecrets2024} uses Shannon entropy to flag
high-randomness strings as likely secrets. The approach casts a wider net than
pure regex matching, at the cost of more false positives.

\textbf{Pre-commit hooks.}
Tools like pre-commit~\cite{precommit2024} plug secret scanners into the git
workflow so that a commit containing a credential is blocked before it reaches
the repository.

\textbf{Cloud-native scanning.}
GitHub Secret Scanning~\cite{github_secret_scanning} and GitLab Secret
Detection~\cite{gitlab_secret_detection} run server-side on push events and
alert maintainers when a known secret format appears.

A shared limitation cuts across all of these approaches: they operate on code
that has already been \emph{committed} or \emph{staged}. None of them address
the scenario in which an AI assistant reads credentials directly from the
working tree, and none can redact credentials from AI conversation transcripts.

\subsection{AI Coding Assistant Security}

Security research around AI coding assistants is still in its early stages, but
several relevant threads have emerged.

\textbf{Prompt injection.}
Greshake et al.~\cite{greshake2023inject} showed that attackers can embed
malicious instructions in repository content (README files, issue templates,
dependency manifests) to hijack an AI assistant's behavior, including
exfiltrating sensitive data.

\textbf{Insecure code generation.}
Pearce et al.~\cite{pearce2022asleep} studied the security quality of code
produced by GitHub Copilot and found that a non-trivial fraction of suggestions
contain vulnerabilities, sometimes including hard-coded credentials lifted from
training data.

\textbf{Data exfiltration through tool use.}
Modern assistants can run shell commands, read arbitrary files, and make network
requests. A manipulated assistant could combine these capabilities to locate and
transmit secrets.

As far as we are aware, VeilGuard is the first tool that sits squarely at the
intersection of credential management and AI coding assistant security.


% ============================================================
\section{Threat Model}
\label{sec:threat}

\subsection{Adversary Model}

We focus on two scenarios.

\textbf{Inadvertent leakage} (the primary threat). A developer's credentials
end up in files that the AI assistant reads: source code, configuration files,
environment files, Model Context Protocol (MCP) configurations, or instruction
files. Once inside the context window, the credential may be sent to a remote
model endpoint, written into a transcript, or repeated back in generated output.

\textbf{Prompt-injection extraction} (secondary). An attacker plants crafted
content in the repository that tricks the AI assistant into reading credential
files, printing environment variables, or smuggling secrets out through
generated code.

\subsection{Trust Boundaries}

Figure~\ref{fig:arch} illustrates VeilGuard's trust boundaries. On the trusted
side sit the local filesystem (protected by POSIX permissions), OS process
isolation, the Python runtime, and the well-audited \texttt{cryptography}
library that provides AES-256-GCM. On the untrusted side are the AI assistant
itself (whose behavior is non-deterministic), the transcript files it writes,
the remote LLM endpoint it communicates with, and repository content that could
carry prompt-injection payloads.

\subsection{Attack Surfaces}

We identify five attack surfaces that are specific to AI coding assistants:

\begin{enumerate}
  \item \textbf{Context window injection.} Credentials embedded in files the AI
        reads (source, config, instruction files, MCP configs).
  \item \textbf{Transcript persistence.} Conversation history written to disk
        by the AI tool; secrets mentioned during a session survive in these
        files.
  \item \textbf{Environment variable exposure.} AI assistants that can run
        shell commands may print or reference environment variables holding
        secrets.
  \item \textbf{Secret store extraction.} The AI may try to read VeilGuard's
        own encrypted store or run credential-extraction commands.
  \item \textbf{MCP configuration exposure.} MCP server configurations
        frequently contain API keys in plain JSON.
\end{enumerate}

\subsection{Out-of-Scope Threats}

VeilGuard does not protect against model-side memorization of training data,
a fully compromised developer machine, interception of the assistant's API
traffic at the network level, or physical access to the key material on disk.


% ============================================================
\section{System Design}
\label{sec:design}

\subsection{Architecture Overview}

VeilGuard is a Python CLI of roughly 2{,}400 lines of source (tests excluded),
organized into eight modules. Figure~\ref{fig:arch} shows how the pieces fit
together.

\begin{figure*}[t]
\centering
\begin{tikzpicture}[
    box/.style={draw, rounded corners=3pt, minimum width=2.4cm, minimum height=1cm,
                font=\small, align=center, thick},
    clibox/.style={draw, rounded corners=3pt, minimum width=2.6cm, minimum height=1.1cm,
                   font=\small, align=center, thick, fill=blue!8},
    arr/.style={-{Stealth[length=3mm]}, thick, black!70},
    darr/.style={{Stealth[length=3mm]}-{Stealth[length=3mm]}, thick, black!70},
    lbl/.style={font=\footnotesize, inner sep=1pt},
  ]

  % --- Row 2 (middle): main components ---
  \node[clibox] (cli) {\textbf{VeilGuard}\\CLI};
  \node[box, fill=red!8,    left=3.5cm of cli]  (ai)    {AI Tool\\(Cursor, Claude, \ldots)};
  \node[box, fill=yellow!8, right=3.5cm of cli] (store) {Encrypted\\Secret Store\\(AES-256-GCM)};

  % --- Row 1 (top): scanner ---
  \node[box, fill=purple!8, above=1.6cm of cli] (scan) {Pattern Scanner\\(55 regexes)};

  % --- Row 3 (bottom): hooks under AI Tool, transcripts under CLI ---
  \node[box, fill=green!8,  below=1.6cm of ai,  minimum width=2.6cm] (hooks) {Hooks / Deny Rules\\/ Instructions};
  \node[box, fill=orange!8, below=1.6cm of cli] (trans) {Transcripts\\(.jsonl)};

  % 1. CLI -> Scanner  (vertical up)
  \draw[arr] (cli) -- node[lbl, right=4pt] {invokes} (scan);

  % 2. Scanner -> AI Tool  (L-shape: left along top, then down)
  \draw[arr] (scan.west) -| node[lbl, above=3pt, pos=0.25] {scans project files} (ai.north);

  % 3. CLI <-> Secret Store  (horizontal)
  \draw[darr] (cli) -- node[lbl, above=3pt] {encrypt / decrypt} (store);

  % 4. CLI -> Hooks  (diagonal down-left)
  \draw[arr] (cli) -- node[lbl, above=3pt, sloped, pos=0.3] {installs} (hooks);

  % 5. Hooks -> AI Tool  (vertical up)
  \draw[arr] (hooks) -- node[lbl, left=4pt] {constrains} (ai);

  % 6. AI Tool -> Transcripts  (diagonal down-right)
  \draw[arr] (ai) -- node[lbl, above=3pt, sloped, pos=0.3] {writes} (trans);

  % 7. CLI -> Transcripts  (vertical down)
  \draw[arr] (cli) -- node[lbl, right=4pt] {clean / watch} (trans);

  % Bounding box
  \begin{scope}[on background layer]
    \node[draw=gray!50, dashed, rounded corners=6pt, line width=1pt,
          fit=(ai)(cli)(store)(hooks)(trans)(scan),
          inner xsep=0.5cm, inner ysep=0.5cm,
          label={[font=\small, fill=white, inner sep=3pt]above:\textit{Developer Machine}}] {};
  \end{scope}
\end{tikzpicture}
\caption{VeilGuard system architecture. The CLI installs hooks and deny rules
into AI tool configurations, invokes the pattern scanner over project files,
redacts transcripts, and manages an encrypted secret store. The dashed
boundary represents the local machine; no data crosses it.}
\label{fig:arch}
\end{figure*}

\subsection{Credential Pattern Engine}
\label{sec:patterns}

At the core of VeilGuard sits a library of 55 compiled regular expressions,
grouped into 8 categories: AI/ML (9 patterns), Cloud (9), Communication (8),
Developer (10), Payment (5), Database (4), Auth (4), and Monitoring (5). Each
pattern is stored as a \texttt{CredentialPattern} dataclass carrying a short
unique ID, a human-readable name, the compiled regex, the conventional
environment variable name, and a category tag.

The patterns are ordered from most specific to least specific, and the scanner
uses a first-match-wins rule on every line. This ordering matters because a
generic pattern like the OpenAI legacy key (\texttt{sk-[a-zA-Z0-9]\{48,\}})
would otherwise shadow more specific prefixes such as
\texttt{sk-ant-api03-} (Anthropic) or \texttt{sk-proj-} (OpenAI project keys).

\textbf{Reducing false positives.} Before the scanner reports a match, three
filters run in sequence:

\begin{enumerate}
  \item A frozen set of \emph{known example keys} drawn from vendor
        documentation (e.g., \texttt{AKIAIOSFODNN7EXAMPLE}).
  \item A list of 21 \emph{placeholder indicators} such as \texttt{example},
        \texttt{fake\_}, \texttt{change\_me}, and \texttt{placeholder}, checked
        against the matched value.
  \item A \emph{comment heuristic} that skips lines whose comment text
        (after \texttt{//} or \texttt{\#}) contains words like ``example'' or
        ``placeholder.''
\end{enumerate}

\textbf{Keeping things fast.} Lines longer than 4{,}096 characters and files
larger than 10\,MB (config files) or 1\,MB (source files) are skipped to cap
regex execution time. A quick-check regex, assembled from the literal prefixes
of all 55 patterns, lets the scanner reject non-matching lines early without
running each pattern individually.

\subsection{AI Tool Integration}
\label{sec:integration}

VeilGuard auto-detects six AI coding tools by looking for filesystem markers
and then installs protections tailored to each one.

\textbf{Claude Code} gets the richest treatment:

\begin{itemize}
  \item A \texttt{PreToolUse} bash hook runs before every file read or shell
        command and blocks access to secret files (\texttt{.env},
        \texttt{*.pem}, \texttt{.aws/credentials}, etc.) as well as commands
        that reference sensitive environment variables.
  \item A \texttt{Stop} hook fires at session end and triggers transcript
        redaction automatically.
  \item Nineteen deny rules in \texttt{settings.json} cover specific
        file-access and command patterns.
  \item Instructions appended to \texttt{CLAUDE.md} tell the model to use
        environment variable references rather than literal credential values.
\end{itemize}

\textbf{Cursor, GitHub Copilot, Windsurf, and Cline} each receive an
instruction-file injection (appended to \texttt{.cursorrules},
\texttt{copilot-instructions.md}, \texttt{.windsurfrules}, or
\texttt{.clinerules}) directing the model away from secret files.

\textbf{Aider} gets file-exclusion patterns in \texttt{.aiderignore} so that
secret files never enter its context.

Table~\ref{tab:tool-integration} summarizes which mechanism applies to which
tool.

\begin{table}[ht]
\centering
\caption{Protection mechanisms available per AI tool.}
\label{tab:tool-integration}
\begin{tabular}{lccccc}
\toprule
Tool & Hook & Deny & Instr. & Ignore & Redact \\
\midrule
Claude Code   & \checkmark & \checkmark & \checkmark &            & \checkmark \\
Cursor        &            &            & \checkmark &            &            \\
Copilot       &            &            & \checkmark &            &            \\
Windsurf      &            &            & \checkmark &            &            \\
Cline         &            &            & \checkmark &            &            \\
Aider         &            &            &            & \checkmark &            \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Transcript Redaction}
\label{sec:transcript}

Claude Code writes conversation history as JSONL files under
\texttt{\textasciitilde/.claude/projects/}. VeilGuard's transcript module
handles these in three steps:

\begin{enumerate}
  \item \textbf{Discovery.} It walks the Claude projects directory recursively,
        collecting \texttt{.jsonl} files and session-memory
        \texttt{summary.md} files.
  \item \textbf{Deep scanning.} Each JSON line is parsed and traversed
        recursively through every nested dictionary, list, and string value. All
        55 patterns are applied at each string leaf. Metadata keys (UUIDs,
        timestamps, session IDs) are skipped so that high-entropy identifiers do
        not cause false matches.
  \item \textbf{Atomic redaction.} Matched credentials are replaced with
        \texttt{[REDACTED:<pattern\_id>]} tokens. Every write goes through a
        tmpfile-rename cycle with \texttt{chmod 0600} to avoid partial writes
        and permission leaks.
\end{enumerate}

A \texttt{watch} mode polls transcripts every 3 seconds, using a PID file to
guarantee only one watcher runs at a time.

\subsection{Encrypted Secret Store}
\label{sec:store}

Instead of hard-coding credentials, developers can store them in VeilGuard's
local encrypted vault. The cryptographic design is straightforward:

\begin{itemize}
  \item \textbf{Key derivation:}
        $\texttt{scrypt}(k, s, n{=}16384, r{=}8, p{=}1) \rightarrow 256\text{-bit key}$,
        where $k$ is a user-supplied passphrase (or a default derived from
        \texttt{HOME} and \texttt{USER}) and $s$ is a 16-byte random salt
        generated once.
  \item \textbf{Encryption:} AES-256-GCM with a fresh 12-byte nonce on every
        write. Ciphertext on disk is simply
        $\text{nonce} \,\|\, \text{AESGCM}(k, n, \text{plaintext})$.
  \item \textbf{File safety:} Advisory locking via \texttt{fcntl.flock},
        atomic tmpfile-rename, and permissions set to \texttt{0600}
        (files) / \texttt{0700} (directories).
\end{itemize}

The store sits behind a \texttt{WritableSecretBackend} protocol, so additional
backends (e.g., cloud KMS adapters) can be plugged in later without touching
the user-facing API.

\subsection{Environment Verification}
\label{sec:verify}

The \texttt{verify} command checks whether any of 45 known credential
environment variables are currently set and then scans both global and
project-level AI context files (instruction files, settings, MCP configs) as
well as the 5 most recent Claude Code transcripts. Verification passes only
when zero credential matches are found in either source.


% ============================================================
\section{Evaluation}
\label{sec:eval}

\subsection{Benchmark Corpus}
\label{sec:corpus}

We assembled a curated benchmark of 93 synthetic files, split into three
groups:

\begin{itemize}
  \item \textbf{58 true-positive files.} One per credential pattern (55
        single-secret files) plus 2 multi-secret files containing 4 credentials
        each and 1 deeply nested JSON file. All credentials are randomly
        generated with realistic formatting and embedded in plausible code
        snippets.
  \item \textbf{25 true-negative files.} Clean code in 13 languages and
        formats. The set includes environment variable references
        (\texttt{\$\{API\_KEY\}}, \texttt{process.env.KEY}), placeholder
        values, known example keys from vendor docs, comments that mention
        credential patterns, and benign high-entropy strings (UUIDs, SHA-256
        hashes, base64 data).
  \item \textbf{10 edge cases.} Lines exceeding 4{,}096 characters, a secret
        sandwiched between safe code, a Redis URI without a password, two
        secrets on one line, a PEM public key header, and a comment next to a
        real credential.
\end{itemize}

A ground-truth manifest maps every file to its expected pattern IDs and line
numbers (or to an expected count for edge cases). The generation script uses a
fixed random seed so that the corpus is fully reproducible.

\subsection{Detection Accuracy}
\label{sec:accuracy}

Table~\ref{tab:detection-accuracy} summarizes VeilGuard's performance on the
benchmark.

\begin{table}[ht]
\centering
\caption{VeilGuard detection accuracy on the benchmark corpus.}
\label{tab:detection-accuracy}
\begin{tabular}{lrrrrrr}
\toprule
 & TP & FP & FN & TN & Prec. & Recall \\
\midrule
Overall & 63 & 0 & 0 & 25 & 1.00 & 1.00 \\
\bottomrule
\end{tabular}
\end{table}

All 63 expected credential instances across the 58 true-positive files are
found, and none of the 25 true-negative files triggers a false alarm. Precision,
recall, and F1 all equal 1.00.

Breaking the numbers down by category confirms clean detection across all 8
groups and all 55 individual patterns. The false-positive filters (known example
keys, placeholder detection, and comment heuristics) correctly suppress
every true-negative file, including those with high-entropy strings, AWS
documentation keys, and credential-like patterns inside comments.

\subsection{Edge Case Analysis}

Of the 10 edge cases, VeilGuard handles each one as expected:

\begin{itemize}
  \item A 4{,}097-character line is skipped (0 findings, as intended).
  \item A credential buried 3 levels deep in a JSON object is detected.
  \item A real credential sitting between two safe \texttt{os.environ}
        references is caught.
  \item A PEM \emph{public} key header does not trip the private-key pattern.
  \item Two secrets on the same line produce one finding (the first-match-wins
        rule is deliberate).
  \item A credential on a line with a ``DO NOT COMMIT'' comment is still
        flagged, because the comment heuristic only fires when the comment
        itself contains a placeholder indicator, not an arbitrary warning.
\end{itemize}

\subsection{Comparison with Existing Tools}
\label{sec:comparison}

We ran three established secret scanners on the same corpus: gitleaks~v8,
trufflehog~v3.94, and detect-secrets~v1.5. Each tool was invoked in its default
filesystem scanning mode against the corpus directory.

\begin{table}[ht]
\centering
\caption{File-level detection comparison on the benchmark corpus.}
\label{tab:tool-comparison}
\begin{tabular}{lrrrrrr}
\toprule
Tool & TP & FP & FN & Prec. & Recall & F1 \\
\midrule
\textbf{VeilGuard} & \textbf{58} & \textbf{0} & \textbf{0} & \textbf{1.00} & \textbf{1.00} & \textbf{1.00} \\
gitleaks       & 43 & 0 & 15 & 1.00 & 0.74 & 0.85 \\
trufflehog     & 21 & 0 & 37 & 1.00 & 0.36 & 0.53 \\
detect-secrets &  0 & 0 & 58 & --   & 0.00 & 0.00 \\
\bottomrule
\end{tabular}
\end{table}

At the file level (Table~\ref{tab:tool-comparison}), VeilGuard catches every
true-positive file. Gitleaks misses 15, mostly AI/ML-specific patterns
(Anthropic, Groq, Perplexity, Fireworks) and newer platform tokens that were
introduced after its rule set was last updated. Trufflehog reaches only 36\%
recall; because it emphasizes server-side verification of detected secrets, many
patterns go unreported when verification is not available. Detect-secrets
returns nothing in a plain filesystem scan because it is designed around a
baseline-diff workflow rather than standalone directory scanning.

Notably, all tools that produce any output achieve 1.00 precision on this
corpus. That result is not surprising: prefix-based credential patterns are
highly specific by nature.

\subsection{Throughput}
\label{sec:throughput}

We measured scan throughput by running VeilGuard 5 times over the full 93-file
corpus (231 total lines) and averaging.

\begin{table}[ht]
\centering
\caption{VeilGuard scan throughput (averaged over 5 runs).}
\label{tab:throughput}
\begin{tabular}{lr}
\toprule
Metric & Value \\
\midrule
Files scanned & 93 \\
Total lines & 231 \\
Mean scan time & 0.010\,s \\
Files / second & 8{,}970 \\
Lines / second & 22{,}282 \\
\bottomrule
\end{tabular}
\end{table}

At roughly 9{,}000 files per second (Table~\ref{tab:throughput}), a full scan
completes in well under a second for any reasonably sized project. For very
large monorepos, a configurable cap (default 5{,}000 source files) prevents
the scan from becoming a bottleneck.


% ============================================================
\section{Discussion}
\label{sec:discussion}

\subsection{Scope and Complementarity}

VeilGuard targets one specific threat class: credentials leaking through AI
coding assistants. It is meant to \emph{complement}, not replace, traditional
secret scanners. A well-rounded security setup would combine VeilGuard for
workstation-level, real-time prevention with gitleaks or trufflehog for
git-history scanning and GitHub Secret Scanning for server-side push protection.

\subsection{Enforcement Strength}

Not all protection mechanisms carry equal weight. Claude Code's
\texttt{PreToolUse} hook is a \emph{hard} gate: the hook script executes before
every file read or shell command, and a deny verdict blocks the operation
outright. Protections for Cursor, Copilot, Windsurf, and Cline depend on the
model obeying instructions in a rules file, which is \emph{advisory} and not
guaranteed. Aider's \texttt{.aiderignore} sits in between: the tool itself
respects the file deterministically, but nothing stops a user from overriding
it.

\subsection{Limitations}

\textbf{Regex coverage.} Pattern matching cannot catch everything. Credential
formats not in the 55-pattern library will slip through, and credentials that
are split across lines, base64-encoded, or stored in binary files will go
undetected.

\textbf{Corpus alignment.} Our benchmark was built hand-in-hand with the
pattern library, so the perfect F1 score reflects that tight alignment more
than it reflects real-world coverage. A future evaluation on a large set of
production repositories would be far more convincing.

\textbf{Transcript support.} Only Claude Code JSONL transcripts are handled
today. Other AI tools may store conversations in proprietary formats that
VeilGuard cannot yet parse.

\textbf{Default key material.} When no custom passphrase is configured, the
encrypted store derives its key from \texttt{HOME} and \texttt{USER}. Any
process running as the same OS user can reproduce these values, so the
protection is limited to cross-user scenarios.

\textbf{POSIX only.} File locking (\texttt{fcntl.flock}) and PID-file
semantics assume a Unix-like environment. Windows is not supported.

\subsection{Threats to Validity}

\textbf{Internal.} The benchmark uses synthetically generated credentials.
Production code may contain credential formats, obfuscation tricks, or
contextual cues that our synthetic samples miss.

\textbf{External.} The tool comparison was run on our own curated corpus, not
on a diverse sample of real-world repositories. Results on production codebases
could differ because of different file-size distributions, language mixes, and
credential format varieties.

\textbf{Construct.} We report file-level detection (whether the tool flags
\emph{at least one} finding in a file) rather than line-level or token-level
metrics. This is the right granularity for the use case (flagging files that
need remediation) but could overstate precision for tools that generate
additional findings within the same file.

\subsection{Ethical Considerations}

Every credential in the benchmark corpus is synthetically generated; no real
secret material was used. VeilGuard itself performs no network communication,
collects no telemetry, and stores no data outside the developer's own machine.
The project is released under the MIT license.


% ============================================================
\section{Implementation Notes}
\label{sec:impl}

VeilGuard requires Python 3.11 or later and has a single runtime dependency:
the \texttt{cryptography} package (for AES-256-GCM). Key numbers:

\begin{itemize}
  \item 2{,}400 lines of source spread across 12 modules.
  \item 140 tests under \texttt{pytest}, covering every module.
  \item Ruff for linting, mypy in strict mode for type checking.
  \item CI on GitHub Actions with a Python 3.11 / 3.12 / 3.13 test matrix.
  \item Published on PyPI as \texttt{veilguard}.
\end{itemize}

The CLI exposes 8 subcommands: \texttt{init} (detect tools, install
protections), \texttt{scan} (credential scan), \texttt{status} (protection
health), \texttt{verify} (environment exposure check), \texttt{clean}
(transcript redaction), \texttt{watch} (continuous transcript monitoring),
\texttt{secret} (store CRUD), and \texttt{backend} (storage backend
management).


% ============================================================
\section{Conclusion}
\label{sec:conclusion}

AI coding assistants introduce a new kind of credential exposure risk: they
read project files and write transcripts, and any secret caught in that flow
can end up on a remote server or on disk in plain text. VeilGuard addresses
this gap with layered defenses covering scanning, prevention, redaction,
verification, and encrypted storage. Our evaluation shows clean detection on a
55-pattern curated corpus and meaningfully better recall than three
established scanners.

Several directions are worth pursuing next: expanding transcript support to
more AI tools, adding entropy-based detection for credential formats that
lack a known prefix, building credential-rotation helpers, and running a
large-scale evaluation on open-source repositories to measure real-world
effectiveness more rigorously.

VeilGuard is available at
\url{https://github.com/manambharadwaj/veilguard} and can be installed with
\texttt{pip install veilguard}.


% ============================================================
\section*{Acknowledgments}

The author thanks the developers behind the Python \texttt{cryptography}
library, Ruff, mypy, and pytest, whose work underpins VeilGuard's
implementation and testing infrastructure.

% ============================================================
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
